Demand Forecasting of individual Probability Density Functions with Machine Learning

### 2020-08-27 LaTex/paper.tex
reworked model and lots of descriptions

\begin{abstract}
Demand forecasting is a central component for many aspects of supply chain operations, as it provides crucial input for subsequent decision making like ordering processes. While machine learning methods can significantly improve prediction accuracy over traditional time series forecasting, the calculated predictions are often mere point estimations for the conditional mean of the underlying probability distribution, and the most powerful approaches, like deep learning, are usually opaque in terms of how its individual predictions can be interpreted. Using the novel supervised machine learning method ``Cyclic Boosting'', complete individual probability density functions in form of negative binomial distributions can be predicted instead of simple point estimates. While metrics evaluating simple point estimates are widely used, methods for assessing the accuracy of predicted distributions are rare and this work proposes new techniques for both qualitative and quantitative evaluation methods. Additionally, each single prediction obtained with this framework is explainable, which is a major benefit in practice, as individual forecasts can be understood by the practitioner and ``black-box'' models can be avoided. 
\end{abstract}

@@

\section{Introduction}

Demand forecasting is one of the main challenges for retailers and at the core of supply chain business operations. Due to its stochastic nature, demand is difficult to forecast. It depends on many influencing factors and the realized demand can be interpreted as a random variable that is described by an appropriate probability density function (PDF). In order to make operational decisions, an optimal point estimator has to be defined, that can be used to derive ordering decisions in the replenishment process of the retailer. Demand estimation is further complicated by the fact that retailers typically only observe realized sales rather than the actual demand, and in case the demand exceeds the current stock level the data become censored.

The ordering decision process is complicated by a range of factors: Even in the case of perfect demand forecasts, the decision maker has to consider lot-sizes defined by the wholesaler or manufacturer as well as to balance conflicting metrics to reach an optimal decision. Ordering too few items may result in out of stock situations leading to unrealized demand and unsatisfied customers. Ordering too many items results in excess inventory, which increases transport and storage costs and, in the case of perishable goods, excessive waste, as spoilt items need to be disposed of at additional cost and potentially even environmental impact. This situation is particularly noticeable in the so called "ultra-fresh'' category, which includes items such as bakery products, ready-meals, fresh diary products, or certain meat products such as ground meat. These items typically have a shelf-life ranging from less than a business day to a few business days at most, with a continuous spectrum in between, depending on the exact item. In many situations, additional constraints have to be considered to reach an optimal ordering decision: Delivery cycles of items may vary depending on the type of item and the wholesaler or manufacturer from which they are procured. Retailers also operate at a given service level to guarantee that a certain level of demand can be fulfilled. The exact service level typically depends on the overall business policy of the retailer and may also depend on individual products, ranging from "never-out-of-stock'' items to a service level exceeding e.g. 90\%.

The availability of Big Data allows capturing, storing, and processing a vast amount of data associated with demand, such as historic sales records, information about promotional events or advertisements, pricing information, local weather at retail locations, seasonal information, as well as a wide range of further variables. Modern machine learning algorithms can then be used to predict the per-item-location demand distribution, corrected for censored data, from which an optimal point estimator can be derived to be used in the subsequent ordering decision. Demand forecasting is an application of time series forecasting and it is important to note that demand as a random variable is not identically and independently distributed (i.i.d.) between different items and locations. While the probability distribution describing the demand can be attributed to a given family or parameterization, the exact parameters vary: Seasonal effects, finite life cycles of products, and the introduction of new products influence the demand distribution, as well as the local weather at the retail location or the retail location itself in terms of size, assortment range, customer diversity, and other factors. The retailers themselves also actively influence demand by using advertisements to highlight products, offering rebates or discounts for specific products as well as pursuing an active pricing strategy. This means that while we can generally assume that demand follows a specific type of probability distribution, its parameters are unique to the instance for which an estimate is required. For example, the probability distribution governing the demand of a particular item is specific to the item, date, and retail location for which the forecast is made, and depends on a wide range of further influencing factors.

The remainder of the paper is organized as follows: We first review the relevant literature and existing work in sec. \ref{sec:LitRev}. We then describe our method to predict individual negative binomial PDFs by means of a parametric approach including two distinct machine learning models for mean of variance in sec. \ref{sec:pdfEstimation}. After that, we describe methods for the qualitative and quantitative evaluation of PDF predictions in sec. \ref{sec:pdfEvaluation}. And finally, we present a demand forecasting example to show an application of our methods in sec. \ref{sec:example}.

@@

Inventory management offers a rich theory and the extensive body of research can be broadly grouped into two categories, where the inventory control problem is either based on some knowledge of the underlying demand distribution or an integrated approach that seeks to map directly from the observed data to the ordering decision. More traditional inventory control systems rely on the knowledge of the demand distribution in one form or another, see e.g. \cite{silver1998} for an overview. In $(s,S)$ type inventory control systems \cite{Scarf1958}, inventory levels are monitored at regular intervals and orders are dispatched once the inventory level reaches a minimal value $s$. In case of linear holding and shortage costs, such policies are optimal \cite{Scarf1959}, although perishable goods pose more challenges, see e.g. \cite{Nahmias1973,Nahmias1975,nahmias1978}. Additionally, service level constraints can be included in these kind of inventory control systems \cite{minner2010periodic}. Perishable goods are well described by the "newsvendor problem'' \cite{Edgeworth}, where, in the simplest case, all stock perishes at the end of the selling period (e.g. a business day). For a detailed review of the newsvendor problem see e.g. \cite{Khouja1999537}. Assuming linear underage and overage costs $b,h >0$, the optimal quantile $q_{\mathrm{opt}} = {b}/{(b+h})$ of a known demand distribution $f(D)$ can be calculated exactly. However, it should be noted that demand is not independent between different samples for a given date. For example, a promotion applied to one product can affect the sales of related products within the assortment. This implies that the demand forecasts cannot be treated as individual newsvendor-type predictions but need to be modeled holistically.

The direct approach is often referred to as "data-driven newsvendor'' and discussed e.g. in \cite{beutel2012safety,ban2019big,bertsimas2020predictive, oroojlooyjadid2020applying}. It aims to avoid estimating the underlying probability distribution for demand and use the available data (historic sales records and further variables) to derive the operational decisions (i.e. the order quantity) directly. An overview of a range of different approaches can also be found in \cite{huber2019data}. Although the integrated approach seems preferable at first glance, since it avoids determining the full demand distribution and results directly in the desired operational decision, the indirect approach via demand forecasts offers some substantial advantages. First, demand forecasts in form of PDFs can be used to simulate the performance on an individual level, and e.g. optimize the impact on business strategy decisions on conflicting metrics such as out of stock (i.e. lost sales) and waste rate. From a practical perspective, separating the demand forecast from the operational decisions (i.e. calculating the order quantities for the next delivery cycle) enables longer-term planning and reduces the complexity, as it avoids coupling delivery schedules of multiple wholesalers and manufacturers with the forecast of customer demand. It also allows to share long-term demand predictions with other business units or external vendors and wholesalers to ease their planning for the production and supply chain processes upstream of the retailer. From the perspective of industrial practice, modeling the demand separately from deriving the subsequent orders has the additional benefit that multiple retail chains can benefit from any improvement in the model description, even if the concrete retailers are unrelated to each other. Additionally, a purely data-driven approach going from the observed data directly to the operational decision (such as the order quantity) does not allow to analyze the data-generating process, i.e. the mechanism behind the stochastic behavior of the customer demand, which is vital if a causal analysis is planned, for example a study of the effect of promotions, advertisements, price changes, or other demand shaping factors in either Pearl's do-calculus \cite{PearlCausality} or Rubin's potential outcomes framework \cite{rubin1974estimating}. Using an operational quantity such as the order quantity will in most cases act as an insufficient proxy for the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that may not be able to be fully controlled for.

So, the main objective in the solution of the inventory control problem is to determine the underlying demand distribution. The simplest approach is to just use the observed sales events and forecast these as a time series (see e.g. \cite{alwan2016}) or via sample average approximation (SAA), see e.g. \cite{shapiro2014} for an overview. However, these approaches do not make use of any data apart from the sales record themselves, although we know that many variables such as price, advertisements, or weather influence, and therefore are highly correlated with, the demand. Saghafian and Tomlin \cite{saghafian2016newsvendor} propose to include partial information about the distribution in the derivation of the operational decision, i.e. the calculation of the optimal order quantity. Often, there are many individual time series, for example millions of item-store combinations to be forecasted daily in retail demand forecasting, and traditional, univariate time series forecasting methods, like autoregressive integrated moving average (ARIMA) \cite{BoxJenkins}, exponential smoothing \cite{ExponentialSmoothing}, or structural time series models \cite{StructuralTS}, operate on each of these time series separately. Machine learning, on the contrary, optimizes all individual time series together, improving the forecast quality of the individual time series by exploiting commonalities between them and therefore reducing the variance of the model. This can be interpreted as kind of hierarchical learning, one example of a hierarchical level being the sum of product sales over all locations in the retail demand forecasting use case. And this reduction of variance does not come with an increase of bias, as usual with the bias-variance tradeoff, because we can still take into account the effects on the finest level. Besides the reduction in variance, this modeling on various hierarchical levels can also be interpreted as kind of transfer learning, improving the generalizability of the method. Furthermore, machine learning enables a natural way to include many exogenous variables, simply as features, what reduces also the bias of the model.

@@

This work demonstrates how the explainable machine learning algorithm Cyclic Boosting \cite{Wick2019} can be used to model the demand distribution at the granularity needed by the retailer. Typically, this means that the full demand PDF has to be estimated per product for each sales location and opening day, conditional on a wide range of variables such as weather, prices, promotions, etc. In contrast to using a "black-box'' machine learning model, Cyclic Boosting allows to interpret each individual prediction in terms of influence of the different features.

@@

To predict an individual PDF using a parametric approach, one has to rely on a model assumption about the underlying distribution of the random variable to be predicted. As discussed earlier, the NBD is well routed in theoretical arguments to model customer demand. Its parameters can be modeled by two independent models, one to estimate the mean and the other for the variance. At least in principle, any method can be used. However, as discussed in sec. \ref{sec:LitRev}, machine learning algorithms are ideally suited for the task of demand forecasting and in the following, we will use the Cyclic Boosting algorithm to benefit in particular from explainable decisions rather than black-box approaches. Furthermore, the regularization approach used during training of the Cyclic Boosting algorithm allows a dedicated treatment of the underlying NBD model, which is another major benefit compared to a standard ''off-the-shelf'' machine learning algorithm. This means we use two subsequent Cyclic Boosting models, in order to estimate the parameters of each individual PDF that we need to forecast, where the first model is used to estimate the mean and the second to estimate the variance, and the features may or may not differ between the mean and variance estimation models. The assigned mean and variance predictions can then be used to generate individual PDFs using the parameterization of the NBD for each sample.

@@

The main idea of this algorithm is the following: First, each feature, denoted by index $j$, is discretized appropriately into $k$ bins to reflect the specific behaviour of the feature. The global mean $\mu$ is determined from all values $y$ of the  target variable $Y \in [0,\infty)$ observed in the data. Single data records, for example the sales of a specific product-location-date combination along with all relevant features, are indexed by $i$.
The individual predictions $\hat{y_i}$  can then be calculated as:
\begin{equation} \label{eqn:cb}
\hat{y}_i = \mu \cdot \prod \limits_{j=1}^p f^k_j \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
\end{equation}
The factors $f^k_j$ are the model parameters that are determined iteratively from the features, and are determined iteratively until the algorithm converges. During training, regularization techniques are applied to avoid overfitting and improve the generalization ability of the algorithm. The deviation of each factor from $f^k_j=1$ can then be used to explain how a  specific feature contributes to each individual prediction.

In detail, the following meta-algorithm describes how the model parameters $f^k_j$ are obtained from the training data:
\begin{enumerate}
\item{Calculate the global average $\mu$ from all observed $y$ across all bins $k$ and features $j$.}
\item{Initialize the factors $f^k_j \leftarrow 1$}
\item{Cyclically iterate through features $j = 1,...,p $ and calculate in turn for each bin $k$ the partial factors $g$ and corresponding aggregated factors $f$, where indices $t$ (current iteration) and $\tau$ (current or preceding iteration) refer to iterations of full feature cycles as the training of the algorithm progresses:
\begin{equation} \label{factors}
g^k_{j,t} = \frac{\sum \limits_{x_{j,i} \in b^k_j} y_i}{\sum \limits_{x_{j,i} \in b^k_j} \hat{y}_{i,\tau}}\;\; \mathrm{where} \; \; f^k_{j,t} = \prod \limits_{s=1}^t g^k_{j,s}
\end{equation}
Here,  $g$ is a factor that is multiplied to $f_{t-1}$ in each iteration. The current prediction, $\hat{y}_\tau$, is calculated according to \eqn \eqref{eqn:cb} with the current values of the aggregated factors $f$:
\begin{equation} \label{factors3}
\hat{y}_{i,\tau} = \mu \cdot \prod \limits_{j=1}^p f^k_{j,\tau}
\end{equation}
To be precise, the determination of $g^k_{j,t}$ for a specific feature $j$ uses $f^k_{j,t-1}$ in the calculation of $\hat{y}$. For the factors of all other features, the newest available values are used, i.e., depending on the sequence of features in the algorithm, either from the current ($\tau=t$) or the preceding iteration ($\tau=t-1$).}
\item{Quit when stopping criteria, e.g. the maximum number of iterations or no further improvement of an error metric such as the mean absolute deviation (MAD) or mean squared error (MSE), are met at the end of a full feature cycle.}
\end{enumerate}

@@

In the previous section, the general Cyclic Boosting algorithm was used to estimate the mean of the NBD model. In order to predict the variance of the NBD model (associated with the mean predicted before), we modify the algorithm like described in the following.

When looking at the demand of individual product-location-date combinations, the target variable $y$ has the values $y = 0, 1, 2, ...$ and the NBD model can be parameterized as in \cite{hilbe2011negative}:
\begin{equation} \label{eqn:nbinom}
\mathrm{NB}(y; \mu, r) = \frac{\Gamma(r + y)}{y! \cdot \Gamma(r)} \cdot \left(\frac{r}{r + \mu}\right)^r \cdot \left(\frac{\mu}{r + \mu}\right)^y,
\end{equation}
where $\mu$ is the mean of the distribution and $r$ a dispersion parameter.

@@

\begin{equation} \label{eqn:r}
\hat{r_i} = 1 + \prod \limits_{j=1}^p f^k_j \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
\end{equation}

@@

The Cyclic Boosting algorithm is trained iteratively using cyclic coordinate descent, processing one feature with all its bins at a time until convergence is reached. Unlike in the basic multiplicative regression mode of Cyclic Boosting described in sec. \ref{sec:CB}, the minimization of the loss function in \eqn \eqref{eqn:loss_likelihood} cannot be solved analytically and has to be done numerically, for example using a random search. All other advantages of Cyclic Boosting, like for example individual explainability of predictions, remain valid for its negative binomial width mode.

@@

\begin{figure}
\begin{center}
\includegraphics[width=2.5cm]{../figures/CDFHistos_uniform}
\includegraphics[width=2.5cm]{../figures/CDFHistos_broad}
\includegraphics[width=2.5cm]{../figures/CDFHistos_narrow}
\includegraphics[width=2.5cm]{../figures/CDFHistos_over}
\includegraphics[width=2.5cm]{../figures/CDFHistos_under}
\caption{\label{fig:cdf_histos} Quantile distribution for different cases of estimating the full probability distribution compared to the expected distribution:
correct prediction (``uniform''), variance overestimated (``broad''), variance underestimated (``narrow''), mean overestimated (``over''), mean underestimated (``under'')}
\end{center}
\end{figure}

@@

We now refine the method from sec. \ref{sec:cdf_histo} by comparing (in the sense of higher or lower) the quantiles of the observed events, i.e. the sales records, with specified quantile values. In order to do so, we start again from the predicted values for the mean and variance from which a negative binomial distribution is constructed for each prediction, for example for the predicted demand for a single product sold on a single day in a given sales location. Each of these predicted negative binomial PDFs is then transformed to its CDF. Note that for simplicity, we always refer to the negative binomial model in this description, however, the general approach is valid for any PDF.

@@

\fig \ref{fig:quant_profiles} illustrates five different collections of quantile profile plots (each collection comparing to 7 specified quantile values), for separate sets of exemplary PDF estimations and observed data combinations. The dashed horizontal lines indicates the fraction we expect, i.e. the specified quantile value, if the predictions are correct. For example, for the median, the line at 0.5 indicates that  50 percent of all PDF prediction and observed data combinations in a given data set should fall above the line, and 50 percent should fall below the line. The observation of the number of samples, indicated with shaded circles, that do in fact fall above and below a particular line, then allows the evaluation of the accuracy of PDF estimations. In case that the PDFs are not estimated correctly, the fractions will deviate from their expected values and the corresponding profile plot allows to judge whether for example the tails of the predicted distribution describing rare events are particularly problematic.

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{../figures/invquant_profiles}
\caption{\label{fig:quant_profiles} Profile plots of quantiles comparing all predicted PDFs to the corresponding observed events. In the leftmost two columns of shaded circles illustrate the behavior if the estimate of the variance is biased, the center and center-right columns of shaded circles illustrate cases for which the mean estimation is biased and the far right column shows the expected behaviour if all predictions are correct.}
\end{center}
\end{figure}

In the same way as the method of filling all observed CDF values of the individual PDF predictions in a histogram and compare to a uniform distribution (described in sec. \ref{sec:cdf_histo}), quantile profile plots do not work on individual PDF predictions but require a statistical population. However, calculating the fractions of the quantile profile plots globally, i.e. over all samples in the data set, might not reveal certain deficiencies for a subset, e.g. a specific store or product group in the example of retail demand forecasts. Therefore, we combine the approach described above with the method of profile plots, where the quantity on the x-axis of the quantile profile plot can be any variable of the data set at hand. Profile plots are akin to scatter plots and described in more detail in appendix \ref{sec:profile}.

>>>>

\begin{abstract}
Demand forecasting is a central component for many aspects of supply chain operations, as it provides crucial input for subsequent decision making like ordering processes. While machine learning methods can significantly improve prediction accuracy over traditional time series forecasting, the calculated predictions are often just point estimations for the conditional mean of the underlying probability distribution, and the most powerful approaches, like deep learning, are usually opaque in terms of how its individual predictions can be interpreted. Using the novel supervised machine learning method ``Cyclic Boosting'', complete individual probability density functions can be predicted instead of single numbers. While metrics evaluating point estimates are widely used, methods for assessing the accuracy of predicted distributions are rare and this work proposes new techniques for both qualitative and quantitative evaluation methods. Additionally, each single prediction obtained with this framework is explainable. This is a major benefit in particular for practitioners, as this allows them to avoid ``black-box'' models and understand the contributing factors for each individual prediction. 
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}

Demand forecasting is one of the main challenges for retailers and is at the core of supply chain business operations. Due to its stochastic nature, demand is difficult to forecast: It depends on many influencing factors and can be interpreted as a random variable that is described by an appropriate probability density function (PDF). Demand estimation is further complicated by the fact that retailers typically only observe realized sales rather than the actual demand, and in case the demand exceeds the current stock level the data become censored. Demand forecasting is an application of time series forecasting and it is important to note that demand as a random variable is not identically and independently distributed (i.i.d.) between different items and locations. While the probability distribution describing the demand can be attributed to a given family or parameterization, the exact parameters vary: Seasonal effects, finite life cycles of products, and the introduction of new products influence the demand distribution, as well as the local weather at the retail location or the retail location itself in terms of size, assortment range, customer diversity, and other factors. The retailers themselves also actively influence demand by using advertisements to highlight products, offering rebates or discounts for specific products as well as pursuing an active pricing strategy. This means that while we can generally assume that demand follows a specific type of probability distribution, its parameters are unique to the instance for which an estimate is required. For example, the probability distribution governing the demand of a particular item is specific to the item, date, and retail location for which the forecast is made, and depends on a wide range of further influencing factors.

Thanks to the availability of Big Data today, a vast amount of these data associated with demand, such as historic sales records, information about promotional events or advertisements, pricing information, local weather at retail locations, seasonal information, as well as many further variables, can be captured, stored, and processed. Different methods, such as traditional univariate time series forecasting or modern supervised machine learning algorithms, can then be used to predict the demand distribution for individual products sold at specific locations, corrected for censored data. Demand forecasting for retailers often requires to predict millions of item-store combinations daily, and traditional univariate time series forecasting methods operate on each of these many individual time series separately. Machine learning, on the contrary, optimizes all individual time series together, improving the generalizability of the method by exploiting common aspects between them. This leads to a drastic reduction in variance of the model, and in turn to an improvement of the forecast quality of each of the individual time series. Furthermore, machine learning enables a natural way to include the many exogenous variables influencing demand as features of the model, what reduces also the bias compared to traditional time series forecasting approaches which mainly rely on the historic sales.

In order to make operational decisions, an optimal demand point estimator has to be defined, that can be used to derive ordering decisions in the replenishment process of the retailer. The ordering decision process is complicated by a range of factors: Even in the case of perfect demand forecasts, the decision maker has to consider lot-sizes defined by the wholesaler or manufacturer as well as to balance conflicting metrics to reach an optimal decision. Ordering too few items may result in out of stock situations leading to unrealized demand and unsatisfied customers. Ordering too many items results in excess inventory, which increases transport and storage costs and, in the case of perishable goods, excessive waste, as spoilt items need to be disposed of at additional cost and potentially even environmental impact. This situation is particularly noticeable in the so called ``ultra-fresh'' category, which includes items such as bakery products, ready-meals, fresh diary products, or certain meat products such as ground meat. These items typically have a shelf-life ranging from less than a business day to a few business days at most, with a continuous spectrum in between, depending on the exact item. In many situations, additional constraints have to be considered to reach an optimal ordering decision: Delivery cycles of items may vary depending on the type of item and the wholesaler or manufacturer from which they are procured. Retailers also operate at a given service level to guarantee that a certain level of demand can be fulfilled. The exact service level typically depends on the overall business policy of the retailer and may also depend on individual products, ranging from ``never-out-of-stock'' items to a service level exceeding, e.g., 90\%. So, demand forecasts are ideally made in the form of full individual probability density distributions, from which optimal point estimators can be derived, which in turn serve as starting point to determine the subsequent ordering decisions.

The remainder of the paper is organized as follows: We first review the relevant literature and existing work in sec. \ref{sec:LitRev}. We then describe our method to predict individual negative binomial PDFs by means of a parametric approach including two distinct machine learning models for mean of variance in sec. \ref{sec:pdfEstimation}. After that, we describe novel techniques for the qualitative and quantitative evaluation of PDF predictions in sec. \ref{sec:pdfEvaluation}. And finally, we present a demand forecasting example to show an application of our methods in sec. \ref{sec:example}.

@@

Inventory management offers a rich theory and the extensive body of research can be broadly grouped into two categories, where the inventory control problem is either based on some knowledge of the underlying demand distribution or an integrated approach that seeks to map directly from the observed data to the ordering decision. More traditional inventory control systems rely on the knowledge of the demand distribution in one form or another, see e.g. \cite{silver1998} for an overview. In $(s,S)$ type inventory control systems \cite{Scarf1958}, inventory levels are monitored at regular intervals and orders are dispatched once the inventory level reaches a minimal value $s$. In case of linear holding and shortage costs, such policies are optimal \cite{Scarf1959}, although perishable goods pose more challenges, see e.g. \cite{Nahmias1973,Nahmias1975,nahmias1978}. Additionally, service level constraints can be included in these kind of inventory control systems \cite{minner2010periodic}. Perishable goods are well described by the ``newsvendor problem'' \cite{Edgeworth}, where, in the simplest case, all stock perishes at the end of the selling period (e.g. a business day). For a detailed review of the newsvendor problem see e.g. \cite{Khouja1999537}. Assuming linear underage and overage costs $b,h >0$, the optimal quantile $q_{\mathrm{opt}} = {b}/{(b+h})$ of a known demand distribution $f(D)$ can be calculated exactly. However, it should be noted that demand is not independent between different samples for a given date. For example, a promotion applied to one product can affect the sales of related products within the assortment. This implies that the demand forecasts for multiple items or a given assortment range cannot be treated as individual newsvendor-type predictions but need to be modeled holistically.

The direct approach is often referred to as ``data-driven newsvendor'' and discussed e.g. in \cite{beutel2012safety,ban2019big,bertsimas2020predictive, oroojlooyjadid2020applying}. It aims to avoid estimating the underlying probability distribution for demand and use the available data (historic sales records and further variables) to derive the operational decisions (i.e., the order quantity) directly. An overview of a range of different approaches can also be found in \cite{huber2019data}. Although the integrated approach seems preferable at first glance, since it avoids determining the full demand distribution and results directly in the desired operational decision, the indirect approach via demand forecasts offers some substantial advantages. First, demand forecasts in form of full PDFs can be used to simulate the performance on an individual level, and, for example, optimize the impact on business strategy decisions on conflicting metrics such as out of stock (i.e., lost sales) and waste rate. From a practitioners perspective, separating the demand forecast from the operational decisions (i.e., calculating the order quantities for the next delivery cycle) enables longer-term planning and reduces the complexity, as it avoids coupling delivery schedules of multiple wholesalers and manufacturers with the forecast of customer demand. It also allows to share long-term demand predictions with other business units or external vendors and wholesalers to ease their planning for the production and supply chain processes upstream of the retailer. From the perspective of the industrial practice of a supplier of supply chain methods and tools, modeling the demand separately from deriving the subsequent orders has the additional benefit that multiple retail chains can benefit from any improvement in the model description, even if the specific retailers are unrelated to each other. Additionally, a purely data-driven approach going from the observed data directly to the operational decision (such as the order quantity) does not allow to analyze the data-generating process, i.e., the mechanism behind the stochastic behavior of the customer demand. This is crucial, for example, if a causal analysis is planned, such as a study of the effect of promotions, advertisements, price changes, or other demand shaping factors in either Pearl's do-calculus \cite{PearlCausality} or Rubin's potential outcomes framework \cite{rubin1974estimating}. Using an operational quantity such as the order quantity will in most cases act as an insufficient proxy for the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that one may not be able to  fully controlled for.

Therefore, the main objective in solving the inventory control problem is to determine the underlying demand distribution. The simplest approach is to use the observed sales records and forecast these via sample average approximation (SAA), see e.g. \cite{shapiro2014} for an overview, or as a time series (see e.g. \cite{alwan2016}), for example by means of autoregressive integrated moving average (ARIMA) \cite{BoxJenkins}, exponential smoothing \cite{ExponentialSmoothing}, or structural time series models \cite{StructuralTS}. However, these approaches do not make use of any data apart from the sales records themselves, although we know that many variables such as price, advertisements, or weather influence, and others are highly correlated with the demand. Saghafian and Tomlin \cite{saghafian2016newsvendor} propose to include partial information about the distribution in the derivation of the operational decision, i.e., the calculation of the optimal order quantity. As described in sec. \ref{sec:intro}, modern supervised machine learning techniques can overcome many of the issues of the more traditional approaches and therefore are the method of choice in most use cases today (see e.g. \cite{tsforecasting,ZHANG199835,rnn,transformer}).

@@

This work demonstrates how the explainable machine learning algorithm Cyclic Boosting \cite{Wick2019} can be used to model the demand distribution at the granularity needed by the retailer. Typically, this means that the full demand PDF has to be estimated per product for each sales location and opening day, conditional on a wide range of variables such as weather, prices, promotions, etc. In contrast to using a ``black-box'' machine learning model, Cyclic Boosting allows to interpret each individual prediction in terms of influence of the different features.

@@

To predict an individual PDF using a parametric approach, one has to rely on a model assumption about the underlying distribution of the random variable to be predicted. As discussed earlier, the NBD is well routed in theoretical arguments to model customer demand. Its parameters can be modeled by two independent models, one to estimate the mean and the other for the variance. At least in principle, any method can be used. However, as discussed in sec. \ref{sec:LitRev}, machine learning algorithms are ideally suited for the task of demand forecasting and in the following, we will use the Cyclic Boosting algorithm to benefit in particular from explainable decisions rather than black-box approaches. Furthermore, the regularization approach used during training of the Cyclic Boosting algorithm allows a dedicated treatment of the underlying NBD model, which is another major benefit compared to a standard ``off-the-shelf'' machine learning algorithm. This means we use two subsequent Cyclic Boosting models, in order to estimate the parameters of each individual PDF that we need to forecast, where the first model is used to estimate the mean and the second to estimate the variance. Hereby, the features may or may not differ between the mean and variance estimation models, and it can be beneficial to include the corresponding mean predictions as feature in the variance model. The assigned mean and variance predictions can then be used to generate individual PDFs using the parameterization of the NBD for each sample.

@@

The main idea of this algorithm is the following: First, each feature, denoted by index $j$, is discretized appropriately into $k$ bins to reflect the specific behaviour of the feature. The global mean $\mu$ is determined from all values $y$ of the  target variable $Y \in [0,\infty)$ observed in the data. Single data records, for example, the sales of a specific product-location-date combination along with all relevant features, are indexed by $i$.
The individual predictions $\hat{y_i}$  can then be calculated as:
\begin{equation} \label{eqn:cb}
\hat{y}_i = \mu \cdot \prod \limits_{j=1}^p f^k_j \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
\end{equation}
The factors $f^k_j$ are the model parameters that are determined iteratively from the features until the algorithm converges. During training, regularization techniques are applied to avoid overfitting and improve the generalization ability of the algorithm. The deviation of each factor from $f^k_j=1$ can then be used to explain how a specific feature contributes to each individual prediction.

In detail, the following meta-algorithm describes how the model parameters $f^k_j$ are obtained from the training data:
\begin{enumerate}
\item{Calculate the global average $\mu$ from all observed $y$ across all bins $k$ and features $j$.}
\item{Initialize the factors $f^k_j \leftarrow 1$}
\item{Cyclically iterate through features $j = 1,...,p $ and calculate in turn for each bin $k$ the partial factors $g$ and corresponding aggregated factors $f$, where indices $t$ (current iteration) and $\tau$ (current or preceding iteration) refer to iterations of full feature cycles as the training of the algorithm progresses:
\begin{equation} \label{factors}
g^k_{j,t} = \frac{\sum \limits_{x_{j,i} \in b^k_j} y_i}{\sum \limits_{x_{j,i} \in b^k_j} \hat{y}_{i,\tau}}\;\; \mathrm{where} \; \; f^k_{j,t} = \prod \limits_{s=1}^t g^k_{j,s}
\end{equation}
Here,  $g$ is a factor that is multiplied to $f_{t-1}$ in each iteration. The current prediction, $\hat{y}_\tau$, is calculated according to \eqn \eqref{eqn:cb} with the current values of the aggregated factors $f$:
\begin{equation} \label{factors3}
\hat{y}_{i,\tau} = \mu \cdot \prod \limits_{j=1}^p f^k_{j,\tau}
\end{equation}
To be precise, the determination of $g^k_{j,t}$ for a specific feature $j$ uses $f^k_{j,t-1}$ in the calculation of $\hat{y}$. For the factors of all other features, the newest available values are used, i.e., depending on the sequence of features in the algorithm, either from the current ($\tau=t$) or the preceding iteration ($\tau=t-1$).}
\item{Quit when stopping criteria, e.g., the maximum number of iterations or no further improvement of an error metric such as the mean absolute deviation (MAD) or mean squared error (MSE), are met at the end of a full feature cycle.}
\end{enumerate}

@@

In the previous section, the general Cyclic Boosting algorithm was used to estimate the mean of the NBD model. In order to predict the variance of the NBD model (associated with the mean predicted before), we modify the algorithm as follows: When looking at the demand of individual product-location-date combinations (meaning the sales record of a specific item sold on a specific day at a specific sales location), the target variable $y$ has the values $y = 0, 1, 2, ...$ and the NBD model can be parameterized as in \cite{hilbe2011negative}:
\begin{equation} \label{eqn:nbinom}
\mathrm{NB}(y; \mu, r) = \frac{\Gamma(r + y)}{y! \cdot \Gamma(r)} \cdot \left(\frac{r}{r + \mu}\right)^r \cdot \left(\frac{\mu}{r + \mu}\right)^y,
\end{equation}
where $\mu$ is the mean of the distribution and $r$ a dispersion parameter.

@@

\begin{equation} \label{eqn:r}
\hat{r_i} = 1 +  \frac{1}{\prod \limits_{j=1}^p f^k_j} \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
\end{equation}

@@

The Cyclic Boosting algorithm is trained iteratively using cyclic coordinate descent, processing one feature with all its bins at a time until convergence is reached. Unlike in the basic multiplicative regression mode of Cyclic Boosting described in sec. \ref{sec:CB}, the minimization of the loss function in \eqn \eqref{eqn:loss_likelihood} cannot be solved analytically and has to be done numerically, for example, using a random search. All other advantages of Cyclic Boosting, like for example individual explainability of predictions, remain valid for its negative binomial width mode.

@@

\begin{figure}
\begin{center}
\includegraphics[width=2.5cm]{../figures/cdf_truth_uniform}
\includegraphics[width=2.5cm]{../figures/cdf_truth_broad}
\includegraphics[width=2.5cm]{../figures/cdf_truth_narrow}
\includegraphics[width=2.5cm]{../figures/cdf_truth_over}
\includegraphics[width=2.5cm]{../figures/cdf_truth_under}
\caption{\label{fig:cdf_histos} Quantile distribution for different cases of estimating the full probability distribution compared to the expected distribution:
correct prediction (``uniform''), variance overestimated (``broad''), variance underestimated (``narrow''), mean overestimated (``over''), mean underestimated (``under'')}
\end{center}
\end{figure}

@@

We now refine the method from sec. \ref{sec:cdf_histo} by comparing (in the sense of higher or lower) the quantiles of the observed events, i.e., the sales records, with specified quantile values. In order to do so, we start again from the predicted values for the mean and variance from which a negative binomial distribution is constructed for each prediction, for example, for the predicted demand for a single product sold on a single day in a given sales location. Each of these predicted negative binomial PDFs is then transformed to its CDF. Note that for simplicity, we always refer to the negative binomial model in this description, however, the general approach is valid for any PDF.

@@

\fig \ref{fig:invquant_example} illustrates five different collections of quantile profile plots (each collection comparing to 6 specified quantile values, namely $q = 0.1$, $q = 0.3$, $q = 0.5$, $q = 0.7$, $q = 0.9$, and $q = 0.97$), for separate sets of exemplary PDF estimations and observed data combinations. The dashed horizontal lines indicates the fraction we expect, i.e., the specified quantile value, if the predictions are correct. For example, for the median, the line at $q = 0.5$ indicates that  50 percent of all PDF prediction and observed data combinations in a given data set should fall above the line, and 50 percent should fall below the line. The observation of the number of samples, indicated with different markers, that do in fact fall above and below a particular line, then allows the evaluation of the accuracy of PDF estimations. In case that the PDFs are not estimated correctly, the fractions will deviate from their expected values and the corresponding profile plot allows to judge whether for example the tails of the predicted distribution describing rare events are particularly problematic.

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{../figures/invquant_example}
\caption{\label{fig:invquant_example} Profile plots of quantiles comparing all predicted PDFs to the corresponding observed events. In the leftmost two columns, the outcome is illustrated if the estimate of the variance is biased, the center and center-right columns illustrate cases for which the mean estimation is biased, and the far right column shows the expected behavior if all predictions are correct.}
\end{center}
\end{figure}

In the same way as the method of filling all observed CDF values of the individual PDF predictions in a histogram and compare to a uniform distribution (described in sec. \ref{sec:cdf_histo}), quantile profile plots do not work on individual PDF predictions but require a statistical population. However, calculating the fractions of the quantile profile plots globally, i.e., over all samples in the data set, might not reveal certain deficiencies for a subset, e.g., a specific store or product group in the example of retail demand forecasts. Therefore, we combine the approach described above with the method of profile plots, where the quantity on the $x$-axis of the quantile profile plot can be any variable of the data set at hand. Profile plots are akin to scatter plots and described in more detail in appendix \ref{sec:profile}.
